---
title: "Problem Set #4"
author: "Chris Onesti"
date: "February 25, 2019"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(gt)
library(ggplot2)
```

## Question 1


```{r echo = FALSE, message = FALSE, warning = FALSE}
poll <- read_csv("ps_4_elections-poll-nc09-3.csv",
            col_types =  cols(
              .default = col_character(),
              turnout_scale = col_double(),
              turnout_score = col_double(),
              w_LV = col_double(),
              w_RV = col_double(),
              final_weight = col_double(),
              timestamp = col_datetime(format = "")))  

q1a_setup <- poll %>% 
  group_by(response) %>% 
  filter(response == "Dem") %>%
  count()

q1a_answer <- q1a_setup$n[1]

q1b_setup <- poll %>% 
  group_by(response) %>% 
  count()

q1b_answer <- q1b_setup$n[3]-q1b_setup$n[4]

q1c_setup <- poll %>%
  filter(gender != gender_combined) %>% 
  count()

q1c_answer <- q1c_setup$n[1]

q1d_setup <- poll %>%
  filter(race_eth == "White", file_race_black != "White") %>% 
  count()


q1d_answer <- q1d_setup$n[1]

q1e_setup_dems <- poll %>% 
  select(response, timestamp) %>% 
  filter(response == "Dem") %>%
  arrange(timestamp)

q1e_setup_reps <- poll %>% 
  select(response, timestamp) %>% 
  filter(response == "Rep") %>% 
  arrange(timestamp)

q1e_answer <- round(as.numeric(q1e_setup_reps$timestamp[1] - q1e_setup_dems$timestamp[1]))


```

There were `r q1a_answer` respondents who supported the Democratic candidate.

There were `r q1b_answer` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (`gender` and `gender_combined`). There are `r q1c_answer` individuals for whom these variables have different values.

There are `r q1d_answer` respondents listed as “White” under `race_eth` who are not listed as “White” under `file_race_black`.

The first `response` of Dem came `r q1e_answer` minutes (rounded to the nearest minute) before the first `response` of Rep.




## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
