---
title: "Problem Set #4"
author: "Chris Onesti"
date: "February 25, 2019"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(gt)
library(ggplot2)
```

## Question 1


```{r echo = FALSE, message = FALSE, warning = FALSE}
#Imports poll data
poll <- read_csv("ps_4_elections-poll-nc09-3.csv",
            col_types =  cols(
              #Sets default column character
              .default = col_character(),
              #Sets turnout scale to double format
              turnout_scale = col_double(),
              #Sets turnout score to double format
              turnout_score = col_double(),
              #Sets w_LV to double format
              w_LV = col_double(),
              #Sets w_RV to double format
              w_RV = col_double(),
              #Sets final weight to double format
              final_weight = col_double(),
              #Set timestamp variable to datetime format
              timestamp = col_datetime(format = "")))  

#Subsets data
q1a_setup <- poll %>%
  #Groups data by response
  group_by(response) %>% 
  #Filters to only Democrat responses
  filter(response == "Dem") %>%
  #Counts responses
  count()

#Sets answer variable
q1a_answer <- q1a_setup$n[1]


#Subsets data
q1b_setup <- poll %>% 
  #Groups by response
  group_by(response) %>% 
  #Counts data
  count()

#Sets answer variable
q1b_answer <- q1b_setup$n[3]-q1b_setup$n[4]

#Subsets data
q1c_setup <- poll %>%
  #Filters data
  filter(gender != gender_combined) %>% 
  #Counts data
  count()

#Sets answer variable
q1c_answer <- q1c_setup$n[1]

#Subsets data
q1d_setup <- poll %>%
  #Filters only values in which the two variables are different
  filter(race_eth == "White", file_race_black != "White") %>% 
  #Counts remaining data
  count()

#Set answer variable
q1d_answer <- q1d_setup$n[1]

#Subsets data
q1e_setup_dems <- poll %>% 
  #selects only response and timestamp variables
  select(response, timestamp) %>% 
  #Filters leaving only Democrat responses
  filter(response == "Dem") %>%
  #Arranges data by time
  arrange(timestamp)

#Subsets data
q1e_setup_reps <- poll %>% 
  #Selects response and timestamp variables
  select(response, timestamp) %>% 
  #Filters leaving only Republican responses
  filter(response == "Rep") %>% 
  #Arrange by timestamp
  arrange(timestamp)

#Sets answer variable as the difference between two times
q1e_answer <- round(as.numeric(q1e_setup_reps$timestamp[1] - q1e_setup_dems$timestamp[1]))


```

There were `r q1a_answer` respondents who supported the Democratic candidate.

There were `r q1b_answer` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (`gender` and `gender_combined`). There are `r q1c_answer` individuals for whom these variables have different values.

There are `r q1d_answer` respondents listed as “White” under `race_eth` who are not listed as “White” under `file_race_black`.

The first `response` of Dem came `r q1e_answer` minutes (rounded to the nearest minute) before the first `response` of Rep.


## Question 2

```{r echo=FALSE}
#Imports data from csv file
orig <- read_csv(file = "ps_4_elections-poll-nc09-3.csv",
                 #Sets default parameters
                 col_types =  cols(
                   #Set default column format to character
                   .default = col_character(),
                   #Set turnout scale to double
                   turnout_scale = col_double(),
                   #Set turnout score to double
                   turnout_score = col_double(),
                   #Set w_LV to double formatted column
                   w_LV = col_double(),
                   #Set w_RV to double formatted column
                   w_RV = col_double(),
                   #Set final weight column to double format
                   final_weight = col_double(),
                   #Set timestamp to date time format
                   timestamp = col_datetime(format = ""))) %>%
  #Select only race, response and weight variables
  select(race_eth, response, final_weight) %>% 
  #Filter leaving only races
  filter(race_eth %in% c("Asian", "Black", "White", "Hispanic", "Other")) %>% 
  #Reorders factors to one shown in example
  mutate(race_eth = fct_relevel(race_eth, c("White", "Black", "Hispanic", "Asian", "Other"))) %>% 
  #Groups data by response and race
  group_by(response, race_eth) %>% 
  #Summarizes data and counts by weight
  summarize(total = sum(final_weight)) %>%
  #Spreads total value across different responses and sets NA values to 0
  spread(key = response, value = total, fill = 0) %>% 
  #Calculates total selected voters
  mutate(all = Rep + Dem + Und + `3`) %>%
  #Converts counts to percentages
  mutate(Dem = Dem / all, Rep = Rep/ all, Und = Und / all) %>%
  #Selects only races and responses in main 3 parties
  select(race_eth, Dem, Rep, Und) %>%
  #Sets 0 values to NA
  na_if(0)

  #Selects data frame
  orig %>% 
    #Creates table
    gt() %>% 
    #Set header
    tab_header(title = "Polling Results in North Carolina 9th Congressional District") %>%
    #Set caption
    tab_source_note("Source: New York Times Upshot/Siena College 2018 live polls") %>%
    #Label the columns
    cols_label(race_eth = "", Dem = "DEM.", Rep = "REP.", Und = "UND.") %>%
    #Format as percent
    fmt_percent(columns = vars(Dem, Rep, Und), decimals = 0) %>%
    #Sets 0 values to NA
    na_if(0) %>%
    #Fill in NA values with dash lines
    fmt_missing(columns = vars(Und), rows = 4) 
  
    #as_raw_html() %>% as.character() %>% cat()
  

```

##Question 3

```{r echo=FALSE}
#Loads poll data
poll %>%
  #Filters out refusals
  filter(educ != "[DO NOT READ] Refused") %>%
  #Reorders factors in education
  mutate(educ = fct_relevel(educ, c("Grade school", "High school", "Some college or trade school", "Bachelors' degree", "Graduate or Professional Degree"))) %>%
  #Creates plot with education and weight
  ggplot(aes(x = educ, y = final_weight)) + 
    #Sets the plot to violin
    geom_violin() + 
    #Flips the coordinates
    coord_flip() +
    #Sets title and subtitle labels
    labs(title = "More Educated Matter Less in North Carolina 9th", subtitle = "Poll gives more weight to people who are less likely to participate in polls", caption = "New York Times Upshot/Siena College 2018 live polls") +
    #Labels the y axis
    ylab("Weight Given to Respondant in Calculating Poll Results") +
    #Labels the x axis
    xlab(NULL) +
    #Adds jitter layer on chart
    geom_jitter(alpha = .4, width = .2)

```

##Question 4

```{r echo=FALSE}
#Stores the number of Republicans
num_rep <- sum(poll$partyid == "Republican")
#Stores the number of Democrats
num_dem <- sum(poll$partyid == "Democrat")
#Reformats response as factors
poll$response <- as.factor(poll$response)
#Reformats partyid as factors
poll$partyid <- as.factor(poll$partyid)

#Starts with data from poll
poll %>% 
  #Select response, partyid, and weights variable
  select(response, partyid, final_weight) %>%
  
  #Reorders data in desired order.  This is done to make the data look cleaner by moving larger parts of the stack to the left while clustering the smaller segments of the stack to the right.
  
  mutate(response = fct_relevel(response, c("Und", "3", "Dem", "Rep"))) %>% 
  #Filters leaving only main 3 parties as identities
  filter(partyid %in% c("Republican", "Democrat", "Independent (No party)")) %>% 
  #Groups data by response and party identification
  group_by(response, partyid) %>%
  #Summarizes data and counts using weights
  summarize(total = sum(final_weight)) %>%
  
  #Plots responses to poll by party identification.  This will be useful to know where the source of each candidates votes are, and will visually demonstrate how many voters are still undecided in each political camp.
  
  ggplot(aes(x = partyid, y = total, fill = response)) + 
  
  #Chose stacked bar plot to display data because it can not only demonstrate the number of registered Republicans, Independents, and Democrats but also their breakdown of responses.
  
  geom_bar(stat = "identity") +
  
  #Changes colors to current political conventions to make it more understandable to someone who is used to seeing many charts like this.
  
  scale_fill_manual(values = c(Rep = "red", Dem = "blue", Und = "gray", `3` = "black")) +
  
  #Flips the coordinates to create a wider figure more suited to screen sizes and webpages 
  
  coord_flip() +
  #Adds the title and subtitle labels
  labs(title = "Small Undecided Independent Vote Could Break Deadlock", subtitle = "Parties retain strong loyalty to candidates.  Independents evenly split.") + 
  #Adds x axis label and sets it to nothing
  xlab(NULL) + 
  #Sets y axis label
  ylab("Vote Count") + 
  #Removes the legend on the side
  theme(legend.title = element_blank())

```

##Colleagues

Chris M., Shafi, Diego M., Debi G., Beau M., others

Used code from class and from "advanced Googling"

Attended Charlie's Study Hall and Dillon's Study Hall

