---
title: "Problem Set #4"
author: "Chris Onesti"
date: "February 25, 2019"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(gt)
library(ggplot2)
```

## Question 1


```{r echo = FALSE, message = FALSE, warning = FALSE}
#Imports poll data
poll <- read_csv("ps_4_elections-poll-nc09-3.csv",
            col_types =  cols(
              .default = col_character(),
              turnout_scale = col_double(),
              turnout_score = col_double(),
              w_LV = col_double(),
              w_RV = col_double(),
              final_weight = col_double(),
              timestamp = col_datetime(format = "")))  

#Subsets data
q1a_setup <- poll %>%
  #Groups data by response
  group_by(response) %>% 
  #Filters to only Democrat responses
  filter(response == "Dem") %>%
  #Counts responses
  count()

#Sets answer variable
q1a_answer <- q1a_setup$n[1]


#Subsets data
q1b_setup <- poll %>% 
  #Groups by response
  group_by(response) %>% 
  #Counts data
  count()

#Sets answer variable
q1b_answer <- q1b_setup$n[3]-q1b_setup$n[4]

#Subsets data
q1c_setup <- poll %>%
  #Filters data
  filter(gender != gender_combined) %>% 
  #Counts data
  count()

#Sets answer variable
q1c_answer <- q1c_setup$n[1]

#Subsets data
q1d_setup <- poll %>%
  #Filters only values in which the two variables are different
  filter(race_eth == "White", file_race_black != "White") %>% 
  #Counts remaining data
  count()

#Set answer variable
q1d_answer <- q1d_setup$n[1]

#Subsets data
q1e_setup_dems <- poll %>% 
  #selects only response and timestamp variables
  select(response, timestamp) %>% 
  #Filters leaving only Democrat responses
  filter(response == "Dem") %>%
  #Arranges data by time
  arrange(timestamp)

#Subsets data
q1e_setup_reps <- poll %>% 
  #Selects response and timestamp variables
  select(response, timestamp) %>% 
  #Filters leaving only Republican responses
  filter(response == "Rep") %>% 
  #Arrange by timestamp
  arrange(timestamp)

#Sets answer variable as the difference between two times
q1e_answer <- round(as.numeric(q1e_setup_reps$timestamp[1] - q1e_setup_dems$timestamp[1]))


```

There were `r q1a_answer` respondents who supported the Democratic candidate.

There were `r q1b_answer` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (`gender` and `gender_combined`). There are `r q1c_answer` individuals for whom these variables have different values.

There are `r q1d_answer` respondents listed as “White” under `race_eth` who are not listed as “White” under `file_race_black`.

The first `response` of Dem came `r q1e_answer` minutes (rounded to the nearest minute) before the first `response` of Rep.


## Question 2

```{r echo=FALSE}
#Imports data from csv file
orig <- read_csv(file = "ps_4_elections-poll-nc09-3.csv",
                 #Sets default parameters
                 col_types =  cols(
                   .default = col_character(),
                   turnout_scale = col_double(),
                   turnout_score = col_double(),
                   w_LV = col_double(),
                   w_RV = col_double(),
                   final_weight = col_double(),
                   timestamp = col_datetime(format = ""))) %>%
  #Select only race, response and weight variables
  select(race_eth, response, final_weight) %>% 
  #Filter leaving only races
  filter(race_eth %in% c("Asian", "Black", "White", "Hispanic", "Other")) %>% 
  #Reorders factors to one shown in example
  mutate(race_eth = fct_relevel(race_eth, c("White", "Black", "Hispanic", "Asian", "Other"))) %>% 
  #Groups data by response and race
  group_by(response, race_eth) %>% 
  #Summarizes data and counts by weight
  summarize(total = sum(final_weight)) %>%
  #Spreads total value across different responses and sets NA values to 0
  spread(key = response, value = total, fill = 0) %>% 
  #Calculates total selected voters
  mutate(all = Rep + Dem + Und + `3`) %>%
  #Converts counts to percentages
  mutate(Dem = Dem / all, Rep = Rep/ all, Und = Und / all) %>%
  #Selects only races and responses in main 3 parties
  select(race_eth, Dem, Rep, Und) %>%
  #Sets 0 values to NA
  na_if(0)

  #Selects data frame
  orig %>% 
    #Creates table
    gt() %>% 
    #Set header
    tab_header(title = "Polling Results in North Carolina 9th Congressional District") %>%
    #Set caption
    tab_source_note("Source: New York Times Upshot/Siena College 2018 live polls") %>%
    #Label the columns
    cols_label(race_eth = "", Dem = "DEM.", Rep = "REP.", Und = "UND.") %>%
    #Format as percent
    fmt_percent(columns = vars(Dem, Rep, Und), decimals = 0) %>%
    #Sets 0 values to NA
    na_if(0) %>%
    #Fill in NA values with dash lines
    fmt_missing(columns = vars(Und), rows = 4) 
  
    #as_raw_html() %>% as.character() %>% cat()


```

##Question 3

```{r echo=FALSE}
#Loads poll data
poll %>%
  #Filters out refusals
  filter(educ != "[DO NOT READ] Refused") %>%
  #Reorders factors in education
  mutate(educ = fct_relevel(educ, c("Grade school", "High school", "Some college or trade school", "Bachelors' degree", "Graduate or Professional Degree"))) %>%
  #Creates plot with education and weight
  ggplot(aes(x = educ, y = final_weight)) + 
    #Sets the plot to violin
    geom_violin() + 
    #Flips the coordinates
    coord_flip() +
    #Sets title and subtitle labels
    labs(title = "More Educated Matter Less in North Carolina 9th", subtitle = "Poll gives more weight to people who are less likely to participate in polls", caption = "New York Times Upshot/Siena College 2018 live polls") +
    #Labels the y axis
    ylab("Weight Given to Respondant in Calculating Poll Results") +
    #Labels the x axis
    xlab(NULL) +
    #Adds jitter layer on chart
    geom_jitter(alpha = .4, width = .2)

```

##Question 4

```{r echo=FALSE}
#Stores the number of Republicans
num_rep <- sum(poll$partyid == "Republican")
#Stores the number of Democrats
num_dem <- sum(poll$partyid == "Democrat")
#Reformats response as factors
poll$response <- as.factor(poll$response)
#Reformats partyid as factors
poll$partyid <- as.factor(poll$partyid)

#Starts with data from poll
poll %>% 
  #Select response, partyid, and weights variable
  select(response, partyid, final_weight) %>%
  #Reorders data in desired order
  mutate(response = fct_relevel(response, c("Und", "3", "Dem", "Rep"))) %>% 
  #
  filter(partyid %in% c("Republican", "Democrat", "Independent (No party)")) %>% 
  group_by(response, partyid) %>%
  summarize(total = sum(final_weight)) %>%
  ggplot(aes(x = partyid, y = total, fill = response)) + geom_bar(stat = "identity") +
  scale_fill_manual(values = c(Rep = "red", Dem = "blue", Und = "gray", `3` = "black")) +
  coord_flip() +
  labs(title = "Small Undecided Independent Vote Could Break Deadlock", subtitle = "Parties retain strong loyalty to candidates.  Independents evenly split.") + 
  xlab("") + 
  ylab("Vote Count") + 
  theme(legend.title = element_blank())


```

##Colleagues

Chris M., Shaffi, Diego M., Debi G., Beau M.

